Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-10-29 bbbq_article.tex
Analyses of chance versus observed epitopes must include uncertainly estimates, fix #233
Improvements by @fransbianchi

Epitope prediction is important to understand the immune system 
and for the design of vaccines.
In this study, we provide evidence that epitopes
derived from TMHs are a major but overlooked source of MHC epitopes. 
Our bioinformatics predictions indicate that TMH-derived epitopes 
are presented to both cytolytic and helper T cells more often than expected by chance, 
regardless of the organism. 
Moreover, reanalysis of peptide elution studies confirmed the presentation of TMH-derived epitopes.
Finally, our SNP analysis shows that TMHs are evolutionary more conserved 
than solvent-exposed protein regions.

@@

% \paragraph{Over-presentation in silico}

An important implication from the over-presentation of TMH-derived epitopes is that T cells will largely recognize 
different protein regions than B cells. 
Presentation of antigens by MHC-II is important 
for the activation of naive B cells by helper T cells. 
For this activation, 
B cells first ingest antigen that is bound to their B cell receptor, 
and subsequently present peptides derived from this antigen in MHC-II 
to helper T cells. 
Following their activation by the T cells, 
B cells mature into plasma cells 
and release antibodies which recognize the same part of the antigen 
as the original B cell receptor. 
B cell receptors and antibodies will thus recognize solvent-exposed regions 
of antigens that are accessible for binding to the B cell receptor. 
However, the results from our study predict that most MHC-II haplotypes 
present relatively hydrophobic peptides,  which are less likely to be solvent-exposed. 
It is unknown why B and T cells seem to predominantly recognize 
different protein regions, but one possibility might be that 
this lowers the chance of B cell mediated autoimmune diseases, 
because auto-reactive B and T cells recognizing different parts 
of the same antigen would need to be present for breakage of B cell tolerance.

@@

  \begin{subfigure}[t]{0.35\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
    \label{fig:hydrophobicity_1}
  \end{subfigure}  
  \hfill
  \begin{subfigure}[t]{0.35\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2.png}
    \label{fig:hydrophobicity_2}
  \end{subfigure}  
  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes}
    \textbf{(A)} 
    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
    row) and \emph{M. tuberculosis} (bottom row).
    The dashed lines indicate the percentages as expected by chance.
    See supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
    for the exact TMH and  epitope counts.
    \textbf{(B-C)}
    Correlation between the percentages of predicted TMH-derived epitopes
    and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
    and MHC-II haplotypes \textbf{(C)}. 
    Diagonal red line: linear regression analysis. 
    Labels are shorthand for the HLA haplotypes,
    see the supplementary Table \ref{tab:haplotype_abbreviations} for the names.
  }
\end{figure}

>>>>

Epitope prediction is important to understand the immune system 
and for the design of vaccines.
In this study, we provide evidence that epitopes
derived from TMHs are a major but overlooked source of MHC epitopes. 
Our bioinformatics predictions indicate that the TMH-derived epitope repertoire is larger than expected by chance, 
regardless of the organism. Furthermore we see despite the bias towards hydrophilic peptides due to their enhanced ionization in mass spec that in several peptide elution studies still a higher fraction of TMH derived epitopes is found than predicted by chance. Therefor it's tempting to assume that this would also result in enhanced T-cell responses, assuming that the conservation of TMH's across organisms does not effect the deletion of such T-cells by central tolerance mechanisms.
Moreover, reanalysis of peptide elution studies confirmed the presentation of TMH-derived epitopes.
Finally, our SNP analysis shows that TMHs are evolutionary more conserved 
than solvent-exposed protein regions.

@@

% \paragraph{Over-presentation in silico}
Although it's well known that T cells and B Cells seem to predominantly recognize different protein regions,
this is enhanced by the over-presentation of TMH-derived epitopes, as TMH's buried in the lipid bilayer are unlikely to be recognized by BCR's.

@@

  \begin{subfigure}[t]{0.35\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
    \label{fig:hydrophobicity_1}
  \end{subfigure}  
  \hfill
  \begin{subfigure}[t]{0.35\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2.png}
    \label{fig:hydrophobicity_2}
  \end{subfigure}  
  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes}
    \textbf{(A)} 
    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
    row) and \emph{M. tuberculosis} (bottom row).
    The pair of dashed lines in each plot indicate the lower and upper bound of the 99\% confidence interval.
    See supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
    for the exact TMH and  epitope counts.
    \textbf{(B-C)}
    Correlation between the percentages of predicted TMH-derived epitopes
    and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
    and MHC-II haplotypes \textbf{(C)}. 
    Diagonal red line: linear regression analysis. 
    Labels are shorthand for the HLA haplotypes,
    see the supplementary Table \ref{tab:haplotype_abbreviations} for the names.
  }
\end{figure}
